---
title: "Final Project - Group 14"
authors: |
    | Andrew Schuck, Charlotte Milard, Hang Tian
    | UB PERSON IDs: ,,50413372
date: "`r Sys.Date()`"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1) # seed for any random generation
```


# 1. "Read data and some base plots"

## Read data and simple processing.
```{r}
suppressPackageStartupMessages(library(tidyverse)) # just in case
library(ISLR2)
library(tidyverse)
library(dplyr)
library(naniar)
library(lubridate)
suppressPackageStartupMessages(library(glmnet)) # penalized linear models
suppressPackageStartupMessages(library(glmnetUtils))  # for quality of life functions over glmnet
suppressPackageStartupMessages(library(corrplot)) # correlation plots
suppressPackageStartupMessages(library(pls)) # for pcr
setwd("~/Semester files/STA 545/STA545_Final_Project")
#call data
origin_data=read_csv('Bike-Sharing-Dataset/hour.csv',show_col_types = FALSE)
#Check how many predictors have NAs
origin_data%>%miss_var_summary()%>%filter(n_miss!=0)%>%nrow()%>%print()
#Avoid changing original data
bs_hour=origin_data%>%mutate(dteday=as.Date(dteday))%>%select(-instant)
#Add one hourly identifiable column to identify every row 
bs_hour=bs_hour%>%mutate(hourly_id=paste(as.character(dteday),as.character(hr)))%>%mutate(hourly_id=ymd_h(hourly_id))
bs_hour=bs_hour[,c(1:15,17,16)]
bs_hour$windspeed=as.numeric(bs_hour$windspeed)
```

## Scatter plots & Box plots for total counts.
```{r,out.width='50%'}
col_vec_scatter=colnames(bs_hour)[10:15]
col_vec_box=colnames(bs_hour)[2:9]
for (value in col_vec_scatter) {
    print(ggplot(bs_hour)+geom_point(aes_string(value,'cnt'))+theme_bw()+
            theme(axis.title.y=element_text(size=16),
        axis.title.x=element_text(size=16)))
}
for (value in col_vec_box) {
    print(ggplot(bs_hour)+geom_boxplot(aes_string(value,'cnt',group=value))+theme_bw())+
            theme(axis.title.y=element_text(size=16),
        axis.title.x=element_text(size=16))
}
cor(bs_hour[, -c(1,16)]) %>% 
  corrplot::corrplot()
```

## Scatter plots & Box plots for casual user counts.
```{r,out.width='50%'}
# col_vec_scatter=colnames(bs_hour)[10:13]
# for (value in col_vec_scatter) {
#     print(ggplot(bs_hour)+geom_point(aes_string(value,'casual'))+theme_bw()+
#             theme(axis.title.y=element_text(size=16),
#         axis.title.x=element_text(size=16)))
# }
# for (value in col_vec_box) {
#     print(ggplot(bs_hour)+geom_boxplot(aes_string(value,'casual',group=value))+theme_bw())+
#             theme(axis.title.y=element_text(size=16),
#         axis.title.x=element_text(size=16))
# }
```

## Scatter plots & Box plots for registered user counts.
```{r,out.width='50%'}
# col_vec_scatter=colnames(bs_hour)[10:13]
# for (value in col_vec_scatter) {
#     print(ggplot(bs_hour)+geom_point(aes_string(value,'registered'))+theme_bw()+
#             theme(axis.title.y=element_text(size=16),
#         axis.title.x=element_text(size=16)))
# }
# for (value in col_vec_box) {
#     print(ggplot(bs_hour)+geom_boxplot(aes_string(value,'registered',group=value))+theme_bw())+
#             theme(axis.title.y=element_text(size=16),
#         axis.title.x=element_text(size=16))
# }
```

# 2. Problems through the data and answering.

## Question 1. Would the hourly distributions of bikeshare users on working days / non-working days different, and what about casual / registered users?
### Total bikeshare users
```{r}
ggplot(bs_hour)+
  geom_boxplot(aes(hr,cnt,group=interaction(workingday,hr),fill=factor(workingday)),outlier.shape = NA)+
  theme_bw()+
  xlab('Hour')+ylab('Hourly records')+
  labs(fill='Workingday',title='Hourly distribution of total bikeshare users')+
  scale_x_continuous(breaks=seq(0,24,2))
```

**The answer is Yes.** \newline
**Before 6am,** there are a few bike share users for both working-day types while more people on non-working days tend to use bike share from 0am to 2am. \newline
**From 6am to 11pm,** Two peaks of users are shown around 8am and 5pm on working days, which may reflect commuting during the rush hours. While on non-working days, we saw a smooth increasing then decreasing trend on bike share users.\newline


### Registered bikeshare users
```{r}
ggplot(bs_hour)+
  geom_boxplot(aes(hr,registered,group=interaction(workingday,hr),fill=factor(workingday)),outlier.shape = NA)+
  theme_bw()+
  xlab('Hour')+ylab('Hourly records')+
  labs(fill='Workingday',title='Hourly distribution of registered bikeshare users')+
  scale_x_continuous(breaks=seq(0,24,2))
```
**The hourly distribution of registered users are quite like that of the total users.**

### Casual bikeshare users
```{r}
ggplot(bs_hour)+
  geom_boxplot(aes(hr,casual,group=interaction(workingday,hr),fill=factor(workingday)),outlier.shape = NA)+
  theme_bw()+
  xlab('Hour')+ylab('Hourly records')+
  labs(fill='Workingday',title='Hourly distribution of casual bikeshare users')+
  scale_x_continuous(breaks=seq(0,24,2))
```
**Casual users tend to use bikeshare more often on non-workding days while there is no strong evidence they would use bikeshare for commuting on rush hours.**


## Question 2. Would bikeshare users counts on different weather be different, and what about casual / registered users?
### Total bikeshare users
```{r}
ggplot(bs_hour)+
  geom_boxplot(aes(weathersit,cnt,group=factor(weathersit),fill=factor(weathersit)),outlier.shape = NA)+
  theme_bw()+
  xlab('Weathersit')+ylab('Hourly records')+
  labs(fill='Weathersit',title='Total bikeshare users on different weather')
```
**Here's the description for weather type 1 - 4:**
weathersit : \newline
- 1: Clear, Few clouds, Partly cloudy, Partly cloudy \newline
- 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist \newline
- 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds \newline
- 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fogv \newline
From these boxplots, it's obvious that from weather type 1 - 4, people are becoming more unwillingly to use bike share. This makes sense because commonly speaking, ordinary people would see weather type from 1 - 4 as weather getting worse. And for type 4, it's definitely bad weather.\newline

### Registered bikeshare users
```{r}
ggplot(bs_hour)+
  geom_boxplot(aes(weathersit,registered,group=factor(weathersit),fill=factor(weathersit)),outlier.shape = NA)+
  theme_bw()+
  xlab('Weathersit')+ylab('Hourly records')+
  labs(fill='Weathersit',title='Registered bikeshare users on different weather')
```
### Casual bikeshare users
```{r}
ggplot(bs_hour)+
  geom_boxplot(aes(weathersit,casual,group=factor(weathersit),fill=factor(weathersit)),outlier.shape = NA)+
  theme_bw()+
  xlab('Weathersit')+ylab('Hourly records')+
  labs(fill='Weathersit',title='Casual bikeshare users on different weather')
```

## Question 3s. Digging deeper into rush hour patterns. Which would affect people's willingness of using bikeshare during rush hours?

## Question 3s - 1. Influence of weathertype
First let's find out those records during rush hours and likely to be considered as commuting.
```{r}
bs_rush=bs_hour%>%filter(hr %in% c(7,8,9,16,17,18,19))%>%filter(workingday==1)
bs_rush$hr=as.factor(bs_rush$hr)
```

### Total bikeshare users
```{r}
ggplot(bs_rush)+
  geom_boxplot(aes(hr,cnt,group=interaction(weathersit,hr),fill=factor(weathersit)),outlier.shape = NA)+
  theme_bw()+
  xlab('Hour')+ylab('Hourly records')+
  labs(fill='Weathersit',title='Hourly distribution of total bikeshare users under different weather')
```
**During the rush hours in the morning**, weather type 1 and 2 seems have similar pattern, a little cloudy weather won't affect people's choice in the morning. Only when weather getting worse to type 3 or 4, people would not willing to ride a bike. \newline
**However during the evening rush hours**, things are different. As the weather getting worse, even a little cloudy would make people not willing to ride a bike. \newline
What about for casual or registered users?

### Registered bikeshare users
```{r}
ggplot(bs_rush)+
  geom_boxplot(aes(hr,registered,group=interaction(weathersit,hr),fill=factor(weathersit)),outlier.shape = NA)+
  theme_bw()+
  xlab('Hour')+ylab('Hourly records')+
  labs(fill='Weathersit',title='Hourly distribution of registered bikeshare users under different weather')
```
**Similar to total users.**

### Casual bikeshare users
```{r}
ggplot(bs_rush)+
  geom_boxplot(aes(hr,casual,group=interaction(weathersit,hr),fill=factor(weathersit)),outlier.shape = NA)+
  theme_bw()+
  xlab('Hour')+ylab('Hourly records')+
  labs(fill='Weathersit',title='Hourly distribution of casual bikeshare users under different weather')
```
**Weather influences are similar to total users. While as shown in previous hourly distribution of casual users, casual users are more willing to use bikehsare in the afternoon.**

## Question 3s - 2. Influence of feeling temperature, humidity and windspeed
In this dataset, all the values of meteorological data are normalized. Which makes us easier to classify them into two levels: high and low (or another medium level).
```{r}
f=function(x,a){
  (max(x)-min(x))*a+min(x)
}
bs_rush$atemp_rank=ifelse(bs_rush$atemp>=f(bs_rush$atemp,0.5),'High','Low')%>%as.factor()
bs_rush$windspeed_rank=ifelse(bs_rush$windspeed>=f(bs_rush$windspeed,0.5),'High','Low')%>%as.factor()
bs_rush$humi_rank=ifelse(bs_rush$hum>=f(bs_rush$hum,0.5),'High','Low')%>%as.factor()
```

### Total bikeshare users
```{r}
ggplot(bs_rush)+
  geom_boxplot(aes(atemp_rank,cnt,
                   group=atemp_rank,fill=factor(atemp_rank)),outlier.shape = NA)+
  theme_bw()+
  xlab('atemp_rank')+ylab('Hourly records')+
  labs(fill='Atemp rank',title='Influence of feeling temperature')#+guides(fill=guide_legend(ncol=2))
```

```{r}
ggplot(bs_rush)+
  geom_boxplot(aes(windspeed_rank,cnt,
                   group=windspeed_rank,fill=factor(windspeed_rank)),outlier.shape = NA)+
  theme_bw()+
  xlab('windspeed_rank')+ylab('Hourly records')+
  labs(fill='windspeed rank',title='Influence of wind speed')#+guides(fill=guide_legend(ncol=2))
```
```{r}
ggplot(bs_rush)+
  geom_boxplot(aes(humi_rank,cnt,
                   group=humi_rank,fill=factor(humi_rank)),outlier.shape = NA)+
  theme_bw()+
  xlab('humi_rank')+ylab('Hourly records')+
  labs(fill='humidity rank',title='Influence of humidity')#+guides(fill=guide_legend(ncol=2))
```

# 3. Random forest regression to predict hourly bikeshare user counts

```{r,out.width='90%'}
library(randomForest)
model_table=bs_rush[,-c(1,6,8,14,15,16)]
str(model_table)
model_table$season=as.factor(model_table$season)
model_table$yr=as.factor(model_table$yr)
model_table$mnth=as.factor(model_table$mnth)
model_table$weekday=as.factor(model_table$weekday)
model_table$weathersit=as.factor(model_table$weathersit)

set.seed(1)
train_rows=sample(nrow(model_table),dim(model_table)[1]*0.7)
train_set=model_table[train_rows,]
test_set=model_table[-train_rows,]
test_x=test_set[,-11]
test_y=test_set[,11]
cnt.rf <- randomForest(cnt ~ ., data = train_set,
                         importance = TRUE,ntree=400,xtest=test_x,ytest=test_y$cnt,
                         keep.forest=TRUE)
print(cnt.rf)
plot(cnt.rf)
y_pred=predict(cnt.rf,test_x)
table=cbind(y_pred,test_y$cnt)%>%as.data.frame()
colnames(table)=c('y_pred','y_true')
ggplot(table)+geom_point(aes(y_pred,y_true))+geom_abline(slope=1,intercept=0,color='red')
```

```{r,out.width='90%'}
#30 most important attributes
varImpPlot(cnt.rf, n.var = min(30, nrow(cnt.rf$importance)),
main = 'Top 30 - variable importance')
```

```{r}
library(nlme)
frm.1 = formula(cnt~season+yr+mnth+hr+weekday+
                  weathersit+temp+atemp+hum+windspeed+
                  atemp_rank+windspeed_rank+humi_rank)
lme.cnt = lme(frm.1,
             control = lmeControl(opt = "optim"),
             random=list(~1|hr),
             data = model_table)
fittedpm <- data.frame(lme.cnt$fitted)
# performance at model fitting : 
result_table=cbind(model_table$cnt,fittedpm$hr)%>%as.data.frame()

thisMSE <- sum((fittedpm[,2]-model_table[,11])^2)/(nrow(fittedpm))
thisRMSE <- sqrt(thisMSE)
thisR2 <- 1 - (sum((model_table[,11]-fittedpm[,2])^2)/
                 sum((model_table[,11]-mean(model_table$cnt))^2))
thisR2
colnames(result_table)=c('y_true','y_pred')
ggplot(result_table)+geom_point(aes(y_pred,y_true))+geom_abline(slope=1,intercept=0,color='red')
```

